---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# taxizedbextra

<!-- badges: start -->

[![Codecov test coverage](https://codecov.io/gh/selkamand/taxizedbextra/branch/master/graph/badge.svg)](https://app.codecov.io/gh/selkamand/taxizedbextra?branch=master)
<!-- badges: end -->

taxizedbextra wraps provides some convenience wrappers and extra utility for taxizedb database. It assumes all taxids (taxonomyIDs) are from the ncbi taxonomy database.

Functions provided are optimized for long vectors with lots of repeated elements.
They also allow users to input taxids <=0 so long as they specify what they're 'scientific names' should be. 


## Installation

You can install the development version of taxizedbextra from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("selkamand/taxizedbextra")
```

## Usage

### Download ncbi taxonomic database

```{r, eval=FALSE}
library(taxizedb)

## download ncbi taxonomy database (just over 2gb) (only run this once)
db_download_ncbi()

# Check where database was downloaded
print(tdb_cache)
```


### Use taxizedbextra functions

```{r example}
library(taxizedbextra)

# convert taxids to names
taxid2name(562)

# convert taxids to names with custom taxid name specification for negative taxids
taxid2name(c(562, -1), special_taxid_names = c("Unknown Microbe" = -1))

# Get lineage string
taxid2lineage(taxids = c(562, 28901))

# Get lineage string with ranks shown
taxid2lineage(taxids = c(562, 28901), show_ranks = TRUE)

# Customise which rank levels are included in lineage
taxid2lineage(taxids = c(562, 28901), ranks_to_include = c("superkingdom","genus", "species"))
```

